Transfer Learning in Visual and Relational Reasoning

### 2020-02-09 Transfer_Learning/cvpr_2nd_submission/experiments_reasoning_transfer.tex
Proofread Reasoning transfer section in Experiments.

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-transfer-clevr}
In the following experiments, we used CoGent-A only.
Using the question groups (categories) defined by the authors, we organized training and testing splits with the goal of measuring whether mastering reasoning for one questions group can help learning others.

@@

We first trained a single SAMNet instance and measured its accuracy on each of the task group $t$ separately.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as "Train on All", serve as baselines for the next two sets of experiments.

We next trained and tested 5 SAMNet instances, each on a single task group $t$ ("Train on t only").
We observe significant accuracy drops (from 18 up to almost 35 points) in all task groups, aside of \textit{Query Attribute}.
We hypothesize this results from the fact that mastering visual grounding and recognizing object attributes (being the core of the latter task) is a necessary skill the system needed by the other tasks.
Thus training on \textit{Query Attribute} helps others, whereas without any training on that particular group the system struggles with developing such a skill.

Finally, we trained 5 SAMNet instances in the extreme reasoning transfer setup: for each task group $t$ we trained a single instance on all tasks but $t$, and next tested its performance on $t$ only.
As expected, this setup appeared to be difficult to deal with, observing severe performance drops, with accuracy of \textit{Count} and \textit{QueryAttribute} dropping to zero.
The explanation is that these categories expect  answers (digits 0-9, names of attributes) that do not overlap with answers in the other groups (binary yes/no), thus the model cannot predict these labels on in a zero-shot transfer.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}
In order to further investigate whether reasoning transfer was effective in leveraging information gained by training a task family at a higher level of the hierarchy we also conducted a setof  experiments using the Canonical variant of COG.
Order of the performed experiments is analogical to the one from previous section.
First column (\cref{fig:cog_reasoning_transfer}, "Train on All") is our baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that those results are in fact weighted averages when aggregating accuracies of our model on the Canonical variant (\cref{fig:samnet_cog_detailed}).

@@

Next, we experimented with the "Train on t only" setup, i.e. training and testing each of 5 SAMNet instances on a single task group $t$ (i.e. single leaf of the proposed hierarchy).
For the \textit{Basic}, \textit{Spatial} and \textit{Obj-Attr} task groups the accuracy is matching the baseline, suggesting that each group contain tasks sufficient for developing all necessary skills.
Therefore models trained on those will not benefit from any additional training on other task groups.
However, the accuracy drops in the models trained only on tasks from the \textit{Compary} and \textit{Cognitive} task groups suggests the oposite.

For that reason we have performed the third set of experiments ("Train on Group A/B"), where we trained two SAMNet instances: first on tasks from the \textit{Group A} aggregate, tested next on each task from the lowest groups (i.e. \textit{Basic}, \textit{Obj-Attr} and \textit{Compare}) separately; and second on tasks from \textit{Group B}, tested separately on \textit{Spatial} and \textit{Cognitive}.
As expected, training on other tasks from group \textit{Group A} enabled the model to significantly improve on \textit{Compare} task (from 58.6\% to 92.3\%).
Similar improvement (from 68.8\% to 97.1\%) was observed in the second model, when trained on \textit{Group B} and tested on \textit{Cognitive}.
Moreover, the achieved accuracy is in fact higher by 1.4 point from the one achieved when training on all tasks (95.7\%).
This suggest that the model might have been struggling with mastering all that one when faced all kind of questions, maybe lacking the capacity.
Finally, there is an interesting effect for \textit{Obj-Attr} and \textit{Spatial}: training on the whole group A/B seems not to help the models, and when comparing those accuracies with the ones achieved by the models trained on task $t$ only we observe some slight drops (around 0.3-0.5 point).
Understanding that effect seems to require further investigation on dependencies between the resoning required by the particular questions and reasoning transfer.

>>>>

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-transfer-clevr}
In the following experiments, we used CoGent-A only.
Using the question groups defined by the authors, we organized training and testing splits with the goal of measuring whether mastering reasoning for one questions group can help learning others.

@@

We first trained a single SAMNet instance on CoGenT-A and measured its accuracy on each of the task group $t$ separately.
Those results, indicated in \cref{fig:cogent_reasoning_transfer} as "Train on All", serve as baselines for the next two sets of experiments.

We next trained and tested 5 SAMNet instances, each on a single task group $t$ ("Train on t only").
We observe significant accuracy drops (from 18 up to almost 35 points) in all task groups, aside of \textit{Query Attribute}.
We hypothesize \textit{Query Attribute} requires mastering visual grounding and recognizing object attributes, as opposed to the other groups which may not enforce it.
Thus, training on \textit{Query Attribute} helps others, whereas without any training on that particular group, SAMNet battles developing such skill.

Finally, we trained 5 SAMNet instances as follows: for each task group $t$ we trained a single instance on all tasks but $t$, and tested its performance on $t$ only.
As expected, this setup appears challenging, causing severe performance drops, with accuracy on \textit{Count} and \textit{QueryAttribute} dropping to zero.
This stems from the non-overlapping labels spaces of \textit{Count} (digits 0 to 9) and \textit{QueryAttribute} (all attributes names) with the other groups (binary "yes" / "no"). Thus, the model cannot predict these labels in a zero-shot transfer.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}
In order to further investigate whether reasoning transfer is effective in leveraging information gained by training a task family at a higher level of the hierarchy, we also conducted a set of experiments using the Canonical variant of COG.
The order of the performed experiments is analogical to the one in the previous section.
The first column (\cref{fig:cog_reasoning_transfer}, "Train on All") is the baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that these results are weighted averages of the accuracies of our model on the Canonical variant tasks (\cref{fig:samnet_cog_detailed}).

@@

Next, we experimented with the "Train on t only" setup, i.e. training and testing each of 5 SAMNet instances on a single task group $t$ (i.e. single leaf of the proposed hierarchy).
For \textit{Basic}, \textit{Spatial} and \textit{Obj-Attr}, the accuracy is matching the baseline ("Train on All"), suggesting that each group contain sufficient information for developing all necessary skills.
Therefore, models trained on these groups may not benefit from any additional training on other task groups.
However, the accuracy drops for \textit{Compary} and \textit{Cognitive}, suggesting the opposite.

Therefore, we have performed a final set of experiments ("Train on Group A/B"), where we trained two SAMNet instances: a) trained on all tasks from \textit{Group A} and tested on each task from the lowest groups (i.e. \textit{Basic}, \textit{Obj-Attr} and \textit{Compare}) separately; b) trained on all tasks from \textit{Group B} and tested separately on \textit{Spatial} and \textit{Cognitive}.
As expected, training on the other tasks from \textit{Group A} enabled the model to significantly improve on the \textit{Compare} task (from 58.6\% to 92.3\%).
A similar improvement (from 68.8\% to 97.1\%) was observed in the second experiment, when trained on \textit{Group B} and tested on \textit{Cognitive}.
Moreover, the achieved accuracy is in fact higher by 1.4 point than the one achieved when training on all tasks (95.7\%).
This suggests the model may have lacked capacity to master questions from all tasks and that training on a smaller subset allowed better performance on this subset.
Finally, we can interestingly note for \textit{Obj-Attr} and \textit{Spatial} that training on the group A/B does not appear to help. When comparing with "Train on t only", we observe slight accuracy drops (around 0.3-0.5 point).
Understanding this phenomenon warrants further investigations on the dependencies between the reasoning required by the particular questions and the transferred reasoning.
